---
title: "Fully Dark-Adapted Fv/Fm Analysis"
subtitle: "From measurements taken by a Walz Diving PAMII"
author: "Sarah Tanja"
date: "`r format(Sys.time(), '%d %B, %Y')`"  
format:
  html:
    df-print: paged
    toc: true
    toc-location: right
    smooth-scroll: true
    link-external-icon: true
    link-external-newwindow: true
    code-fold: false
    code-tools: true
    code-copy: true
    highlight-style: breeze
    code-overflow: wrap
    theme: minty
editor: 
  markdown: 
    wrap: 72
---

# Overview

Generalized Linear Mixed Models (GLMM) resources:

Mixed Design Examples:
<https://courses.washington.edu/psy524a/_book/linear-mixed-models.html#using-lmer-for-a-repeated-measures-design:~:text=example%2C%20mixed%20designs%3A-,24.3%20Mixed%20Design%20Example%3A%20Effect%20of%20Napping%20and%20Time%20on%20Perceptual%20Performance,-A%20mixed%2Ddesign>

> This is a mixed design: The experiment has one within-subjects factor
> and two between-subjects factors. The days of exposure are the
> within-subjects factor (each coral was measured four times), and the
> leachate condition (control vs. low vs. high) and the temperature
> (ambient vs. hot) are the between-subjects factors.

# Install and load packages

```{r}
# Install packages
if ("tidyverse" %in% rownames(installed.packages()) == 'FALSE') install.packages('tidyverse')
if ("dplyr" %in% rownames(installed.packages()) == 'FALSE') install.packages('dplyr')
if ("ggplot2" %in% rownames(installed.packages()) == 'FALSE') install.packages('ggplot2')
if ("hrbrthemes" %in% rownames(installed.packages()) == 'FALSE') install.packages('hrbrthemes')
if ("lme4" %in% rownames(installed.packages()) == 'FALSE') install.packages('lme4')
if ("lmerTest" %in% rownames(installed.packages()) == 'FALSE') install.packages('lmerTest')
if ("Rmisc" %in% rownames(installed.packages()) == 'FALSE') install.packages('Rmisc')


# Load packages
library(dplyr)
library(tidyverse)
library(ggplot2)
library(hrbrthemes)
library(lme4)
library(lmerTest)
library(Rmisc)
library(RColorBrewer)
```

# Pull in Fv/Fm data

```{r}
fvfm_alldays <- read_csv("../output/fvfm_alldays.csv")
```

Let's do an initial exploration of the data. We should have Fo, Fm, and
FvFm as the measuring variables as well as sample_id, temp, colony,
leachate, days_of_exposure, and treatment as the explanatory variables.
The treatment column is a concatenation of the temp and leachate groups.

```{r}
summary(fvfm_alldays)
```

# Is it normal?

```{r}
# get the means and standard errors
fvfm.summary <- summarySE(fvfm_alldays, measurevar="FvFm", groupvars=c("leachate","temp","days_of_exposure"))
```

# Graphing color palette

```{r}
display.brewer.all()
```

```{r}
# Select specific colors
color_heat <- brewer.pal(9, "YlOrBr")[5]  
# Select the 5th color (a shade of orange/red)
color_ambient <- brewer.pal(9, "YlGnBu")[5]  
# Select the 5th color (a shade of blue/green) from "BuGn"

# Print the selected colors
print(color_heat)
print(color_ambient)
```

```{r}
# Set the order of the temp factor
fvfm_alldays$temp <- factor(fvfm_alldays$temp, levels = c("hot", "ambient"))

# Set the order of the leachate factor
fvfm_alldays$leachate <- factor(fvfm_alldays$leachate, levels = c("control", "low", "high"))

# Define custom colors for temp using ColorBrewer
temp_colors <- c("hot" = brewer.pal(9, "YlOrBr")[7], 
                      # Orange/Red for heat
                   "ambient" = brewer.pal(9, "YlGnBu")[5]) 
                      # Blue/Green for ambient
```

# Line plot

```{r}
ggplot(fvfm.summary, aes(x=days_of_exposure, y=FvFm, color = temp, shape = leachate)) +
  geom_errorbar(aes(ymin=FvFm-se, ymax=FvFm+se), width=.2) +
  geom_line() +
  theme_bw() +
  geom_point(size = 2.5) +
  scale_color_manual(values = temp_colors) # Custom colors for temp
```

# Boxplot

fvfm_alldays is the dataframe Create the boxplot with custom colors for
temp and greyscale for leachate concentration

## By Day

```{r}
fvfm_facet <- ggplot(fvfm_alldays, aes(x = leachate, y = FvFm, color = temp, fill = leachate)) +
  geom_boxplot() +
  facet_grid(. ~ days_of_exposure, switch = "x") +
  theme_bw() +
  labs(
    title = "Photosynthetic efficiency (Fv/Fm) across Treatments and Days of Exposure",
    x = "Days of Exposure",
    y = "Photosynthetic efficiency (Fv/Fm)"
  ) +
  theme(axis.text.x = element_blank()) +  
    # Remove x-axis labels
  scale_color_manual(values = temp_colors) +  
    # Custom colors for temp
  scale_fill_brewer(palette = "Greys")  
    # Pale sequential fill for leachate

fvfm_facet

# save the plot as a .png
ggsave("../output/fvfm_facet_plot.png", plot = fvfm_facet, width = 10, height = 6, dpi = 600)
```

## By Treatment

```{r}
# Create the boxplot
fvfm_byday_treatment <- ggplot(fvfm_alldays, aes(x = as.factor(treatment), y = FvFm, color = as.factor(days_of_exposure))) +
  geom_boxplot() +
  facet_grid(. ~ treatment, switch = "x", scales = "free_x") +
  theme_bw() +
  labs(
    title = "Photosynthetic efficiency (Fv/Fm) across Treatments and Days of Exposure",
    x = "Treatment",
    y = "Photosynthetic efficiency (Fv/Fm)"
  ) +
  theme(axis.text.x = element_blank(), # remove x-axis labels
    strip.text = element_text(size = 6)  # adjust facet label size
  )

fvfm_byday_treatment

# save the plot as a .png
ggsave("../output/fvfm_byday_treatment_plot.png", plot = fvfm_byday_treatment, width = 10, height = 6, dpi = 600)
```

```{r}
# Create a new column combining days_of_exposure and treatment
fvfm_alldays$day_treatment <- interaction(fvfm_alldays$days_of_exposure, fvfm_alldays$treatment)

# Set the order of the leachate factor
fvfm_alldays$day_treatment <- factor(fvfm_alldays$day_treatment, 
                                     levels = c("control", "low", "high"))

# Create the boxplot
fvfm_combo <- ggplot(fvfm_alldays, aes(x = as.factor(day_treatment), y = FvFm, fill = leachate, color = temp)) +
  geom_boxplot() +
  theme_bw() +
  labs(
    title = "Photosynthetic efficiency (Fv/Fm) across Treatments and Days",
    x = "Days of Exposure and Treatment",
    y = "Photosynthetic efficiency (Fv/Fm)"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
    scale_color_manual(values = temp_colors) +  
    # Custom colors for temp
   scale_fill_brewer(palette = "Greys")  
    # Pale sequential fill for leachate

fvfm_combo

# save the plot as a .png
ggsave("../output/fvfm_combo_plot.png", plot = fvfm_combo, width = 10, height = 6, dpi = 600)

```

# GLMM

```{r}
lmer.fvfm <- lmer(FvFm ~ leachate*temp*days_of_exposure + (1|sample_id), data=fvfm_alldays)

anova(lmer.fvfm)
```

> ::: callout-important
> Duration of experiment had a significant effect (p = 0.02158) \* on
> dark-adapted quantum yield (Fv/Fm)
> :::
